---
title: "Assignment5"
---


===============================================================================================================================

**Task1**

Explore all accounts. Create visualizations that combine or these account characteristics:
 - Whether an account has a credit card or not
 - Whether an account has a loan or not
 - The average balance for the account
```{r setup, include=FALSE}
library(tidyverse)
```
```{r,echo=F}
options( warn = -1 )
df = read.csv('Assignment5/accounts_analytical.csv')
task1 = df[,c('account_id','credit_cards','loan_amount',"max_balance","min_balance")]
task1$credit_cards[is.na(task1$credit_cards)] <- 0
task1$loan_amount[!is.na(task1$loan_amount)] <- T
task1$loan_amount[is.na(task1$loan_amount)] <- F
colnames(task1)[3] <- 'loan'
task1$loan <- as.factor(task1$loan)
task1$avg_balance <- (task1$max_balance+task1$min_balance)/2

task1 %>% mutate(cc = ifelse(credit_cards==0, FALSE, TRUE)) %>%
  ggplot(aes(x=loan, y=avg_balance)) +
  geom_point(aes(color=cc, shape=loan,alpha=0.3),position=position_jitter(0.4)) +
    labs(x = "loan status",
           y = "average balance", title = "Loan vs Number of credit cards vs Average balance")

```

===============================================================================================================================

**Task2**

What is the distribution of all loans and what are their characteristics?


```{r, echo=F}
df %>% select(c('loan_amount')) %>% drop_na(.) %>% 
  ggplot(aes(x=loan_amount))+
  geom_histogram(bins = 50)+
  ggtitle('loan_amoount distribution')
```
From above graph, the distribution of loan amount is skew to right. Most loans are less than 500.

```{r, echo=F}

na.rm <- function(x){gsub("^$", NA, trimws(x))}
df$loan_date = sapply(df$loan_date,na.rm)

df %>% select(c('loan_date')) %>% 
  drop_na(.) %>% 
  mutate(loan_year_month = substring(loan_date,0,7)) %>%
  mutate(loan_year = substring(loan_date,0,4)) %>%
  ggplot(aes(x=loan_year_month,fill = loan_year))+
  geom_bar(stat = "count")+
  ggtitle('loan_amoount distribution')+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
From above graph, the amount of loans for each year was increasing in year 95,96, and 97.

===============================================================================================================================

**Task3**

There is no apparent relationship bewteen loan default and the difference between account creatation date and loan apply date.
```{r, echo=F}
df %>% mutate(date.diff = as.integer(as.Date(loan_date,"%Y-%m-%d")-as.Date(acct_creation_date,"%Y-%m-%d"))) %>%
  drop_na(loan_default) %>%
  ggplot(aes(x=loan_default,y=date.diff))+
  geom_boxplot()

```

Certain districts tend to have higher default rate.
```{r, echo=F}
df %>%
  drop_na(account_district) %>%
  drop_na(loan_default) %>%
  group_by(account_district) %>%
  mutate(default_rate = mean(loan_default)) %>%
  ggplot(aes(x = reorder(account_district,-default_rate), fill = loan_default)) +
    geom_bar(stat = "count")+
    theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```




================================================================================================================================

**Task4**

```{r, echo=F}
df1 = read.csv('Assignment5/transactions.csv')
max_acct = names(which.max(table(df1$account_id)))

df1 %>% filter(account_id == max_acct) %>%
  mutate(time_range = as.Date(date)) %>%
  arrange(time_range) %>%
  ggplot(aes(x=time_range,y=amount,color=type)) +
  geom_line() +
  scale_x_date(date_labels = "%Y %b %d")

df1 %>% filter(account_id == max_acct) %>%
  mutate(method = as.character(method)) %>%
  filter(method %in% c("bank transfer", "cash", "credit card")) %>%
  mutate(time_range = as.Date(date)) %>%
  arrange(time_range) %>%
  ggplot(aes(x=time_range,y=amount,color=method)) +
  geom_line() +
  scale_x_date(date_labels = "%Y %b %d")

df1 %>% filter(account_id == max_acct) %>%
   mutate(time_range = as.Date(date)) %>%
  arrange(time_range) %>%
  ggplot(aes(x=time_range,y=amount,color=category)) +
  geom_line() +
  scale_x_date(date_labels = "%Y %b %d")
```

================================================================================================================================

**Task5**

```{r, echo=F}
df$credit_cards = ifelse(is.na(df$credit_cards),F,T)
df1 = merge(x = df1, y = df[,c("account_id","credit_cards")], by = "account_id", all.x = TRUE)
df1[,c("account_id","method","credit_cards")] %>% 
  filter(as.character(method) == "credit card") %>%
  distinct() %>%
  ggplot(aes(x=credit_cards,y=account_id)) +
  geom_point(shape="-",size=30, alpha=0.05)
```
