---
title: "Final Project Network"
output: 
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
library(jsonlite)
library(tidyverse)
library(shiny)
library(stringr)
library(stringi)
library(visNetwork)
options( warn = -1 )

json_file = 'data/events/events_England.json'
events = fromJSON(json_file,flatten = T)

#test = head(events,10000)
#write_json(test, "test.json")

#json_file = 'test.json'
events = fromJSON(json_file,flatten = T)
events = rename(events,eventwyid = id)
events <- events %>% 
  select(-c(eventId,subEventId))
```


### Passing Network

This is all the pass events recorded in a game that Mancester City absolutely destroyed their opponent, Swansea City. The result is 5 - 0. From the passing network, we can tell Man City maintained strong mid-field control and forward offensiveness. **Feel free to move thing around and come up with your formation of the game.** Go to the link [here](https://www.premierleague.com/match/22688) and click Line-ups to check out the true answer!


```{r}
#2500065 1625
match.id = 2500065
#match.id = 2499719
team.id = 1625 #1609
### specific match pass events
df <- events %>%
  filter(matchId == match.id & teamId == team.id & eventName=="Pass") %>%
  select(c("positions", "playerId", "eventwyid")) %>%
  unnest(positions) %>%
  filter(row_number() %% 2 == 1) %>%
  mutate(from1 = ifelse(row_number() %% 2==1, "from", "to"))%>%
  mutate(from2 = ifelse(row_number() %% 2==1, "to", "from"))

### player info
json_file = 'data/players.json'
players = fromJSON(json_file,flatten = T)

### pass from
from1 <- df %>%
  filter(from1=="from") %>%
  left_join(players[c("wyId", "shortName", "role.name")], by=c("playerId" = "wyId")) %>%
  rename(from_y = y, from_x = x, from_id = playerId,
         from_role = role.name, from_who = shortName) %>%
  select(-c("from1","from2")) 
from1$from_who = stri_unescape_unicode(gsub("<U\\+(....)>", "\\\\u\\1", from1$from_who)) 


a <- nrow(df)-1

from2 <- df[2:a,] %>%
  filter(from2=="from") %>%
  left_join(players[c("wyId", "shortName", "role.name")], by=c("playerId" = "wyId")) %>%
  rename(from_y = y, from_x = x, from_id = playerId,
         from_role = role.name, from_who = shortName) %>%
  select(-c("from2","from1")) 
from2$from_who = stri_unescape_unicode(gsub("<U\\+(....)>", "\\\\u\\1", from2$from_who)) 

from <- rbind(from1,from2)

### pass to
to1 <- df%>%
  filter(from1=="to") %>%
  left_join(players[c("wyId", "shortName", "role.name")], by=c("playerId" = "wyId")) %>%
  rename(to_y = y, to_x = x, to_id = playerId,
         to_role = role.name, to_who = shortName) %>%
  select(-c("from1","from2"))
to1$to_who = stri_unescape_unicode(gsub("<U\\+(....)>", "\\\\u\\1", to1$to_who))

a <- nrow(df)-1

to2 <- df[2:a,]%>%
  filter(from2=="to") %>%
  left_join(players[c("wyId", "shortName", "role.name")], by=c("playerId" = "wyId")) %>%
  rename(to_y = y, to_x = x, to_id = playerId,
         to_role = role.name, to_who = shortName) %>%
  select(-c("from1","from2"))
to2$to_who = stri_unescape_unicode(gsub("<U\\+(....)>", "\\\\u\\1", to2$to_who))

to <- rbind(to1,to2)
```

```{r,fig.width=10,fig.height=7}
### list of edges
df <- cbind(from,to)

### build up network

# nodes
nodes <- rbind(from[c("from_id","from_who", "from_role")], setNames(to[c("to_id", "to_who", "to_role")], c("from_id","from_who", "from_role")))

nodes <- nodes %>%
  distinct(from_id,from_who,from_role) %>%
  rename(id = from_id, label = from_who, group = from_role)

# edges
edges <- df %>%  
  select(-c("eventwyid")) %>%
  group_by(from_id, to_id) %>%
  summarise(weight = n()) %>% 
  ungroup() %>%
  rename(from = from_id, to = to_id)

edges <- mutate(edges, width = weight/5 + 1)




visNetwork(nodes, edges) %>% 
  visIgraphLayout(layout = "layout_with_fr") %>% 
  visEdges(arrows = "middle") %>%
  visGroups(groupname = "Forward", color = "red", shape = "triangle", 
            shadow = list(enabled = TRUE)) %>% 
  visGroups(groupname = "Midfielder", color = "lightblue", shape = "circle", 
            shadow = list(enabled = TRUE)) %>%
  visGroups(groupname = "Defender", color = "green", shape = "shield", 
            shadow = list(enabled = TRUE)) %>%
  visGroups(groupname = "Goalkeeper", color = "yellow", shape = "star", 
            shadow = list(enabled = TRUE)) %>%
    visLegend()



```
















